### 2019-03-19 2_metg.tex
read metg

Minimum effective task granularity (METG) is defined intuitively as
the smallest average task granularity such that the application
achieves overall efficiency exceeding some threshold. For example, in
a compute-bound application efficiency might be measured as the
percentage of the available FLOPS achieved. On a machine with a peak
capability of 1.2 TFLOPS per node, METG(50\%) would correspond to
smallest task granularity achieved while maintaing at least 0.6
TFLOPS. Absolute efficiency (i.e., FLOPS or bytes per second) is
preferrable, but parallel efficiency (relative to a sufficiently large
problem size) can also be used in applications which are not amenable
to being characterized by peak resource utilization.

@@

To calculate METG, this data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOPS achieved) and task
granularity (i.e., $\text{wall time} \times \text{number of
  cores}/\text{number of tasks}$). Note that a \emph{task} is defined
broadly to be any continuously-executing unit of work in the
application, thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI codes written in
a bulk synchronous style.

Figure~\ref{fig:efficiency-mpi}
shows the replotted data from above. In this chart we can see that
efficiency starts near 100\%, and that initially task granularity
drops quickly with minimal change in efficiency. However as task
granularity shrinks, efficiency begins to drop as well, leading
eventually to a plateau in task granularity. Intuitively, this plateau
represents the fundamental baseline overhead of the system, the point
beyond which the system simply cannot execute tasks efficiently.

@@

METG is an appealing metric in part because it has a well-defined
relationship with quantities of interest to application developers,
namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the behavior of MPI weak and strong
scaling, respectively, at a variety of problem sizes. Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figurs, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to show up at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed and begin to trend upward. At the
very bottom, the lines compress together as it the fundamental
overhead of MPI dwarfs any gains from running smaller
problems. Critically, the shape of the curve at the bottom is the same
for both weak and strong scaling; this is the precisely the shape of
the METG curve.

@@

METG therefore has a direct relationship with the smallest problem
size that can be weak scaled to a given node count with a given level
of efficiency. Any smaller and the overhead begins to dominate useful
computation. Similarly, METG corresponds to the point at which strong
scaling can be expected to stop; as strong scaling runs start with a
larger problem size and progressively shrink it until overheads begin
to dwarf the gains due to continued scaling. In
Figures~\ref{fig:weak-scaling-mpi} and \ref{fig:strong-scaling-mpi},
the dashed, bright red line show the value of METG(50\%) at each node
count.

>>>>

Minimum effective task granularity (METG) is defined intuitively as
the smallest average task granularity such that the application
achieves overall efficiency exceeding some threshold. For example, in
a compute-bound application efficiency might be measured as the
percentage of the available FLOPS achieved. On a machine with a peak
capability of 1.2 TFLOPS per node, METG(50\%) would correspond to
smallest task granularity achieved while maintaining at least 0.6
TFLOPS. Absolute efficiency (i.e., FLOPS or bytes per second) is
preferable, but parallel efficiency (relative to a sufficiently large
problem size) can also be used in applications which are not amenable
to be characterized by peak resource utilization.

@@

To calculate METG, this data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOPS achieved) and task
granularity (i.e., $\text{wall time} \times \text{number of
  cores}/\text{number of tasks}$). Note that a \emph{task} is defined
broadly to be any continuously-executing unit of work in the
application, thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI codes are written in
a bulk synchronous style.

Figure~\ref{fig:efficiency-mpi}
shows the replotted data from above. In this chart, we can see that
efficiency starts near 100\%, and that initially task granularity
drops quickly with a minimal change in efficiency. However as task
granularity shrinks, efficiency begins to drop as well, leading
eventually to a plateau in task granularity. Intuitively, this plateau
represents the fundamental baseline overhead of the system, the point
below which the system simply cannot execute tasks efficiently.

@@

METG is an appealing metric in part because it has a well-defined
relationship with quantities of interest to application developers,
namely weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the behavior of MPI weak and strong
scaling, respectively, at a variety of problem sizes. Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to show up at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed and begin to trend upward. At the
very bottom, the lines compress together as it the fundamental
overhead of MPI dwarfs any gains from running smaller
problems. Critically, the shape of the curve at the bottom is the same
for both weak and strong scaling; this is precisely the shape of
the METG curve.

@@

METG therefore has a direct relationship with the smallest problem
size that can be weak scaled to a given node count with a given level
of efficiency. With any smaller problem size, 
the overhead begins to dominate useful
computation. Similarly, METG corresponds to the point at which strong
scaling can be expected to stop; as strong scaling runs start with a
larger problem size and progressively shrink it until overheads begin
to dwarf the gains due to continued scaling. In
Figures~\ref{fig:weak-scaling-mpi} and \ref{fig:strong-scaling-mpi},
the dashed, bright red line show the value of METG(50\%) at each node
count.
